Continous time Markov chain models; \\ extension, simulation and visualisation

### 2015-02-11 chapters/chapter01.tex
more revision editing

\noindent
where $\mathbf{\Theta}$ represents the tree topology, the branch lengths, the parameters of the substitution model and other parameters of interest that we want to estimate.
This quantity expresses how likely it is to observe the data given a specific value of the parameter(s) of interest $\mathbf{\Theta}$.
\myedit{sveight}{
By finding the value of $\mathbf{\Theta}$ that maximizes (\ref{eq:likelihood}) we obtain an estimate of the most plausible 
%hypothesis for the data.
set of parameters given observed data.}

@@

\myedit{svnine}{
The prior choice is perhaps one of the most controversial points of the Bayesian inference.
The prior distributions may come from the preceding analysis of other sets of data, or they may form some set of biologically plausible values.
However in some cases there may be no suitable objective distribution at hand, in which case one often resorts to the so-called \emph{vague} or \emph{diffuse} priors, which have large variances.
Some popular uninformative priors on continuous, unbounded variables may be improper, \textit{i.e.} their distribution is infinitesimal on an infinite range in order to integrate to $1$.
Among examples of such priors is the uniform distribution on an infinite interval.
Although many of of these prior choices sill lead to a proper posterior distribution, yet some controversies about their choice may still arise.
}

@@

Figure~\ref{fig:bayes1} shows the resulting posterior distribution along with prior and scaled likelihood.
The likelihood of the binomial model used in this example is simply:

@@

Equation~\ref{eq:binomial_like} is plotted in Figure~\ref{fig:bayes1} over a range of values of $\theta$, resulting in the black filled distribution.
The Beta distribution we chose for the prior has a following distribution function:

@@

Equation~\ref{eq:beta_dist} corresponds to the white filled curve in in Figure~\ref{fig:bayes1}.
Following the Bayes rule from Equation~\ref{eq:bayesian1} we can combine the likelihood and prior to form the posterior, which then becomes:

@@

Equation~\ref{eq:beta_binom_post} appears as a grey filled area in Figure~\ref{fig:bayes1}.
This illustrates that the data `pulls' the posterior away from the prior, and the strength by which is does so depends on its variance, resulting in a posterior centered around 0.06 in this case.
In this example, the posterior distribution can be obtained analytically because the beta distribution is a so-called \emph{conjugate prior} for the binomial model, resulting in the posterior that is also beta-distributed.

>>>>

\noindent
where $\mathbf{\Theta}$ represents the tree topology, the branch lengths, the parameters of the substitution model and other parameters of interest that we want to estimate.
This quantity expresses how likely it is to observe the data given a specific value of the parameter(s) of interest $\mathbf{\Theta}$.
\myedit{sveight}{
By finding the value of $\mathbf{\Theta}$ that maximizes (\ref{eq:likelihood}) we obtain an estimate of the most plausible 
%hypothesis for the data.
set of parameters given the observed data.}

@@

\myedit{svnine}{
The prior choice is perhaps one of the most controversial aspects of the Bayesian inference.
The prior distributions may be inspired by previous analyses of other sets of data, or they may form some set of biologically plausible values.
However in some cases there may be no suitable objective prior knowledge at hand, in which case one often resorts to the so-called \emph{vague} or \emph{diffuse} priors, which have large variances.
Some popular uninformative priors on continuous, unbounded variables may be improper, \textit{i.e.} their distribution is infinitesimal on an infinite range in order to integrate to $1$.
Among examples of such priors is the uniform distribution on an infinite interval.
Although many of these prior choices still lead to a proper posterior distribution, some controversies about their choice may still arise, in particular when the data is poorly informative.
}

@@

Figure~\ref{fig:bayes1} shows the resulting posterior distribution along with the prior and the scaled likelihood.
The likelihood of the binomial model used in this example is simply:

@@

Equation~\ref{eq:binomial_like} is plotted in Figure~\ref{fig:bayes1} over a range of values of $\theta$, resulting in the black filled distribution.
The Beta distribution we choose for the prior has a following distribution function:

@@

Equation~\ref{eq:beta_dist} corresponds to the white filled curve in in Figure~\ref{fig:bayes1}.
Following the Bayes rule from Equation~\ref{eq:bayesian1}, we can combine the likelihood and prior to arrive at the posterior, which then becomes:

@@

Equation~\ref{eq:beta_binom_post} appears as a grey filled area in Figure~\ref{fig:bayes1}.
This illustrates that the data `pulls' the posterior away from the prior, and the strength by which is does so depends on its variance, resulting in a posterior centred around 0.06 in this case.
In this example, the posterior distribution can be obtained analytically because the beta distribution is a so-called \emph{conjugate prior} for the binomial model, resulting in the posterior that is also beta-distributed.
